\name: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction

### 2021-03-15 paper/main.tex
Update results section for DTA methods

\title[short Title]{\name: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction}git
\author[Sample \textit{et~al}.]{Tilman Hinnerichs\,$^{\text{\sfb 1,}*}$ and Robert Hoehndorf\,$^{\text{\sfb 2}}$}
\address{$^{\text{\sf 1}}$Department, Institution, City, Post Code, Country and \\
$^{\text{\sf 2}}$Department, Institution, City, Post Code,
Country.}

@@

\begin{table}[!tpb]
	\begin{tabular}{|p{2.2cm}|p{0.9cm}|p{0.75cm}|p{0.9cm}|p{0.75cm}|}
		\hline
		STITCH results&\multicolumn{4}{c|}{PPI graph}\\
		&\multicolumn{2}{c|}{without}&\multicolumn{2}{c|}{with}\\
		&Macro AUC&Micro $AUC_p$&Macro AUC&Micro $AUC_p$\\
		\hline
		MolPred&$0.69$&$0.65$&$0.69$&$0.67$\\
		\hline
		OntoPred&$0.88$&$0.87$&$0.92$&$0.93$\\
		\hline
		MolPred + OntoPred & $0.89$ & $0.90$&$0.93$&$0.94$\\
		\hline
		\hline
		Yamanishi results&\multicolumn{4}{c|}{PPI graph}\\
		&\multicolumn{2}{c|}{without}&\multicolumn{2}{c|}{with}\\
		&Macro AUC&Micro AUC&Macro AUC&Micro $AUC_p$\\
		\hline
		MolPred + OntoPred & $0.83$ & $0.82$&$0.84$&$0.84$\\
		\hline
	\end{tabular} \medskip

	\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
		\hline
		Approach&\multicolumn{2}{c|}{Original scheme}&\multicolumn{2}{c|}{Protein split}\\
		&Splitting scheme&Macro AUC&Macro AUC&Micro $AUC_p$\\
		\hline
		DTINet&DP pairs&$0.91$&$0.84$&$0.67$\\
		DTIGEMS+&DP pairs&$0.93$& $0.72$& $0.68$ \\
		DTI-CDF&Proteins&$0.86$&$0.85$&$0.79$\\
		\hline
		\hline
		Approach&\multicolumn{2}{c|}{Original scheme}&\multicolumn{2}{c|}{Protein split}\\
		&Splitting scheme&Macro AUC&Macro AUC&Micro $AUC_p$\\
		\hline
		DeepDTI&Drugs&$0.88$&$0.76$&$0.70$\\
		DeepDTA&DP pairs&$0.88$&$0.77$&$0.69$\\
		DeepConv-DTI&DP pairs&$0.88$&$0.76$&$0.73$\\
		MolTrans&DP pairs&$0.90$&$0.77$&$0.74$\\
		\hline
	\end{tabular}\\

	\caption{Results for \name on STITCH and Yamanishi dataset evaluated with a 5-fold cross-validation. MolPred and OntoPred are the predictors for molecular and ontology based features, respectively. We hereby denote the molecular feature based predictor with \textit{MolPred}, while abbreviating the ontology based, top-down predictor with \textit{OntoPred}.}
	\label{tab:STITCHresults}
\end{table}

@@

We trained, validated and finally tested all considered models on the
STITCH dataset using a 5-fold cross-validation over a protein
split; we then selected the best-performing models (with respect to
\textit{AUROC/MacroAUC} and $MicroAUC_p$) and evaluate them on in 5-fold protein split cross validation on the Yamanishi benchmark dataset to avoid validation overfitting and yield more realistic testing results. To evaluate the influence of the different
features separately, and determine whether they ``localize'' on the
PPI graph (and therefore can be exploited successfully by the graph
neural networks), we train and evaluate models with different types of
features, and with and without inclusion of the PPI graph,
separately, comparing the molecular (MolPred) and phenotype based (OntoPred) predictor and a combination of both, concatenating both features.
Table \ref{tab:STITCHresults} shows the results of these experiments.

@@

The na\"ive predictor shows great performance on Yamanishi dataset, with substantial gain in comparison to a random predictor on both datasets. In the following we will denote this nai\"ve predictor as a baseline predictor correlating its performance to state of the art models and \name.

@@

@@

\todo[inline]{Replace/merge table:}
\begin{figure}[!tpb]%figure1

	\centerline{}
	\caption{Test}
	\label{tab:Results}
\end{figure}

@@

\todo[inline]{Likely discussion:}
Assuming a hypothetical, perfectly generalizing model built upon an
unbiased dataset, this very model would yield similar performances for
all three splitting schemes, not overfitting on the known
structures. On the other hand, for a hypothetical entirely overfitting
model, trained on a highly biased dataset, this model would show
substantial deviations from the original performance over another
split.

>>>>

\title[short Title]{\name: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction}
\author[Sample \textit{et~al}.]{Tilman Hinnerichs\,$^{\text{\sfb 1,}*}$ and Robert Hoehndorf\,$^{\text{\sfb 2}}$}
\address{$^{\text{\sf 1}}$Department, Institution, City, Post Code, Country and \\
$^{\text{\sf 2}}$Department, Institution, City, Post Code,
Country.}

@@

\begin{table}[!tpb]
	\begin{tabular}{|p{2.2cm}|p{0.9cm}|p{0.75cm}|p{0.9cm}|p{0.75cm}|}
		\hline
		(a) STITCH results&\multicolumn{4}{c|}{PPI graph}\\
		&\multicolumn{2}{c|}{without}&\multicolumn{2}{c|}{with}\\
		&Macro AUC&Micro $AUC_p$&Macro AUC&Micro $AUC_p$\\
		\hline
		MolPred&$0.69$&$0.65$&$0.69$&$0.67$\\
		\hline
		OntoPred&$0.88$&$0.87$&$0.92$&$0.93$\\
		\hline
		\name (MolPred + OntoPred) & $0.89$ & $0.90$&$\mathbf{0.93}$&$\mathbf{0.94}$\\
		\hline
		\hline
		(b) Yamanishi results&\multicolumn{4}{c|}{PPI graph}\\
		&\multicolumn{2}{c|}{without}&\multicolumn{2}{c|}{with}\\
		&Macro AUC&Micro AUC&Macro AUC&Micro $AUC_p$\\
		\hline
		\name (MolPred + OntoPred) & $0.83$ & $0.82$&$0.84$&$0.84$\\
		\hline
	\end{tabular} \medskip

	\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
		\hline
		(c) Approach&\multicolumn{2}{c|}{Original scheme}&\multicolumn{2}{c|}{Protein split}\\
		&Splitting scheme&Macro AUC&Macro AUC&Micro $AUC_p$\\
		\hline
		Naive predictor&DP pairs&$0.85$& --&--\\
		DTINet&DP pairs&$0.91$&$0.74$&$0.67$\\
		DTIGEMS+&DP pairs&$\mathbf{0.93}$& $0.72$& $0.68$ \\
		DTI-CDF&Proteins&$0.85$&$\mathbf{0.85}$&$0.79$\\
		\name&Proteins&$0.84$&$0.84$&$\mathbf{0.84}$\\
		\hline
		\hline
		(d) Approach&\multicolumn{2}{c|}{Original scheme}&\multicolumn{2}{c|}{Protein split}\\
		&Splitting scheme&Macro AUC&Macro AUC&Micro $AUC_p$\\
		\hline
		DeepDTI&Drugs&$0.88$&$0.76$&$0.70$\\
		DeepDTA&DP pairs&$0.88$&$0.77$&$0.69$\\
		DeepConv-DTI&DP pairs&$0.88$&$0.76$&$0.73$\\
		MolTrans&DP pairs&$0.90$&$0.77$&$0.74$\\
		\hline
	\end{tabular}\\

	\caption{(a) + (b) Results for \name on STITCH and Yamanishi dataset evaluated with a 5-fold cross-validation. We hereby denote the molecular feature based predictor with \textit{MolPred}, while abbreviating the ontology based, top-down predictor with \textit{OntoPred}. (c) + (d) Results for various state of the art (c) drug--target interaction prediction methods on Yamanishi dataset and (d) drug--target affinity prediction methods on BIOSNAP dataset, evaluated on their original and the protein cross-validation splitting scheme, approximately reproducing the results of MolTrans \textbf{CITATION}.}
	\label{tab:Results}
\end{table}

@@

We trained, validated and finally tested all considered models on the
STITCH dataset using a 5-fold cross-validation over a protein
split; we then selected the best-performing models (with respect to
\textit{AUROC/MacroAUC} and $MicroAUC_p$) and evaluate them on in 5-fold protein split cross validation on the Yamanishi benchmark dataset to avoid validation overfitting and yield more realistic testing results. To evaluate the influence of the different
features separately, and determine whether they ``localize'' on the
PPI graph (and therefore can be exploited successfully by the graph
neural networks), we train and evaluate models with different types of
features, and with and without inclusion of the PPI graph,
separately, comparing the molecular (MolPred) and phenotype based (OntoPred) predictor and a combination of both, concatenating both features.
Table \ref{tab:Results} (a)+(b) shows the results of these experiments.

@@

The na\"ive predictor shows great performance on Yamanishi dataset, with substantial gain in comparison to a random predictor on both datasets. In the following, we will utilize this nai\"ve predictor as baseline predictor correlating its performance to state of the art models and \name.

@@

The results of this analysis are summarized in part (c) of Table \ref{tab:Results}, calculating the performance of all considered methods over their original splitting scheme and over a protein split. We find that, first the majority of approaches evaluate over a drug--protein splitting scheme, while second, the highest performances on Yamanishi dataset are yielded over the drug--target pairs splitting scheme, with a noticeable gap to the methods performing their splits over proteins. Further, when evaluating the same methods over a protein split, we find a substantial performance divergence in comparison to their original splitting scheme. The authors of DTI-CDF evaluate their method on all three splitting schemes underlining this point. While \name provides comparable performance to DTI-CDF, it yields considerably better with respect to $MicroAUC_p$.

As we were quite surprised about the MacroAUC discrepancy between the splitting schemes, we evaluated other cutting edge drug--target interaction and drug--target affinity prediction methods, training and evaluating on other datasets. Following the results of MolTrans (\textbf{CITATION}), we reevaluated DeepDTI (\textbf{CITATION}), DeepDTA (\textbf{CITATION}), DeepConv-DTI (\textbf{CITATION}) and MolTrans itself on BioSnap (\textbf{CITATION}) over a protein split, as shown in part (d) of Table \ref{tab:Results}. The authors of MolTrans evaluated all considered methods over the drug--target pair and the protein split, which we were able to reproduce with (d), showing once more a substantial divergence. However, we additionally computed the $MicroAUC_p$ score for all considered methods, leading to comparable scores as for the Yamanishi evaluated methods.

@@

\todo[inline]{Replace/merge table:}

@@

\todo[inline]{Likely discussion:}
Assuming a hypothetical, perfectly generalizing model built upon an
unbiased dataset, this very model would yield similar performances for
all three , not overfitting on the known
structures. On the other hand, for a hypothetical entirely overfitting
model, trained on a highly biased dataset, this model would show
substantial deviations from the original performance over another
split.
